
\chapter{Architectural View Decomposition}
\label{ch_architectural_view}

A system must be viewed from different perspectives in order to model, implement and document that system. Therefore the 4+1 Model View will be used to represent the SAFEcrypto software system: Use Case View, Logical View, Development View, Process View and Physical View.


\begin{itemize}
	\item Use Case View: the main purpose of the Use Case View is to define the system requirements.
	\item Logical View: this view contains system definitions and class diagrams to describe the services that the system will provide to its users.
	\item Development View: this view will provide system and user interface specficiations
	\item Process View: this view will use collaberation, sequence and activity diagrams to describe the processes that form the system.
	\item Physical View: this view will depict how the systems' hardware nodes will interact and how they are installed.
\end{itemize}


\section{Stakeholders}
\label{sec:stakeholders}

The following is a list of stakeholders who are actively involved with the project, are affected by the outcome or have an influence in the development of the SAFEcrypto library.

\begin{itemize}
	\item \textbf{Developer} - responsible for the development of the software.
	\item \textbf{Maintainer} - responsible for enhancement and fixing of the software.
	\item \textbf{Researcher} - utilise the project to implement, analyse and evaluate the cryptographic schemes that they develop.
	\item \textbf{Tester} - responsible for quality assessment of the library to ensure it is robust
	\item \textbf{User} - will use the software directly.
\end{itemize}
