\chapter{Introduction}
\label{ch_introduction}

This Introduction provides and overview of the entire \textit{Software Architecture Document} for the SAFEcrypto software system.
It includes the purpose, scope, definitions, acronyms, abbreviations, references and overview of the system.

\section{Purpose}
SAFEcrypto will provide a new generation of practical, robust, and physically secure post-quantum cryptographic functions. A suite of software routines will be developed to implement the lattice-based constructions identified in WP4 of the SAFEcrypto project.

This document provides an architectural overview of the SAFEcrypto software system and all associated deliverables. A number of architectural views are used to describe the software system and provide an overview of its functionality and structure. It is intended that this document will capture and convey the significant architectural decisions which have been made on the software system.

\section{Scope}
This scope of this document includes the entire SAFEcrypto software system including portability and scalability of the codebase, modular code design, dynamic configuration, build infrastructure, testing strategy and benchmarking.


\section{Definitions, Acronyms and Abbreviations}
\begin{labeling}{SAFEcrypto}
\item [SAD] Software Architecture Document
\item [SAFEcrypto] Safe Architectures of Future Emerging Cryptography
\item [WP\textit{N}] Work Package \textit{N}
\end{labeling}

\section{References}
The following documents should be read in conjunction with this \textit{Software Architecture Document}:

\begin{itemize}
\item SAFEcrypto: Secure Architectures of Future Emerging Cryptography, H2020-ICT-2014--1 \cite{safecrypto}
\item SAFEcrypto: Software Requirements Specification, Version 1.0, April 22 2016 (see~\cite{safecrypto_srs})
\item SAFEcrypto: Secure Coding Guidelines, Version 1.0, September 2016 (see~\cite{safecrypto_secure_coding_guidelines})
\end{itemize}

\section{Overview}
This document consists of 9 sections which are described below:

\begin{itemize}
\item Section 1 is simply an introduction the software arachitecture of the SAFEcrypto software system.
\item Section 2 identifies the goals and constraints of the system's architecture.
\item Section 3 describes the architectural representation of the system and shows the five views in which the system documentation is divided by following the \textit{Rational Unified Process}.
\item Section 4 talks about the secure coding guidelines.
\item Section 5 discusses the functional and software security testing.
\item Section 6 provides a description of the size and performance of the software system.
\item Section 7 describes some system quality issues.
\item Section 8 is a bibliography of the references used to create this document.
\item Section 9 provides an Appendices providing discussions on proposed features and implementation details.
\end{itemize}

